---
title: "Individual Analysis Chapter 11"
author: "Denis Kozhokar"
date: "11/17/2018"
output: html_document
---

The dataset looks at factors that influence the decision of whether to apply to graduate school. College juniors are asked if they are unlikely, somewhat likely, or very likely to apply to graduate school. Therefore, the ouctome variable has 3 different levels. Also, there is data on parental educational status, whether the undergraduate institution is public or private, and current GPA. The is a strong reason to believe that the distances between the 3 different levels.

```{r}
library(foreign)
library(MASS)
library(Hmisc)
library(reshape2)
library(rethinking)
dat <- read.dta("https://stats.idre.ucla.edu/stat/data/ologit.dta")
head(dat)

#descriptives
lapply(dat[, c("apply", "pared", "public")], table)

#3-way cross-table
ftable(xtabs(~ public + apply + pared, data = dat))
```



```{r}
as.character(dat$apply)
dat$apply <- as.numeric(dat$apply)
simplehist( dat$apply , xlim=c(1,3) , xlab="response" )

lapply(dat[, c("apply", "pared", "public")], table)

# discrete proportion of each response value
pr_k <- table( dat$apply ) / nrow(dat)
# cumsum converts to cumulative proportions
cum_pr_k <- cumsum( pr_k )

# plot
plot( 1:3 , cum_pr_k , type="b" , xlab="response" ,
ylab="cumulative proportion" , ylim=c(0,1) )

```

#log-cumulative-odds that a response value yi is equal-to-or-less-than some possible outcome
```{r}

# convenience function
logit <- function(x) log(x/(1-x))
( lco <- logit( cum_pr_k ) )

```

#Fitting a basic model by incorporating no predictor variables:
```{r}
m11.1 <- map(
    alist(
        apply ~ dordlogit( phi , c(a1,a2) ),
        phi <- 0,
        c(a1,a2) ~ dnorm(0,10)
),
data=dat , start=list(a1=-2,a2=-1) )

precis(m11.1)

#cumalitive probabilites
logistic(coef(m11.1))

```



#Run an MCMC
```{r}
m11.1stan <- map2stan(
    alist(
        apply ~ dordlogit( phi , cutpoints ),
        phi <- 0,
        cutpoints ~ dnorm(0,10)
),
data=list(apply=dat$apply), start=list(cutpoints=c(-2,-1)) , chains=2 , cores=2 )
# need depth=2 to show vector of parameters
precis(m11.1stan,depth=2)
```


#Model Comparison
```{r}
m11.2 <- map(
    alist(
        apply ~ dordlogit( phi , c(a1,a2) ) ,
        phi <- bPA*pared + bPu*public + bGPA*gpa,
         c(bPA,bPu,bGPA) ~ dnorm(0,10),
          c(a1,a2) ~ dnorm(0,10)
        ) ,
        data=dat ,
        start=list(a1=-1.9,a2=-1.2) )
precis(m11.2)

m11.3 <- map(
    alist(
        apply ~ dordlogit( phi , c(a1,a2) ) ,
        phi <- bPA*pared + bPu*public + bGPA*gpa + 
          bPAPu*pared*public + bGPAPu*gpa*public,
         c(bPA,bPu,bGPA,bPAPu,bGPAPu) ~ dnorm(0,10),
          c(a1,a2) ~ dnorm(0,10)
        ) ,
        data=dat ,
        start=list(a1=-1.9,a2=-1.2) )

precis(m11.3)        
```

```{r}
coeftab(m11.1,m11.2,m11.3)
compare( m11.1 , m11.2 , m11.3 , refresh=0.1 )

```
#It seems like b11.2 is the best model based on it having the lowest WAIC. b11.3 has the interaction terms but still has a lower WAIC. b11.2 will be used further on and for interpretation.

```{r}
precis(m11.2)
```

# Public demonstrates a negative effect on the expected value of applying. The confidence metric crosses 0. Therefore, there is not strong effect as the effect can be positive or negative. For every one unit increase in pared (from 0-1), there is a 1.05 increase in the expected value of apply on the log odds scale, given all of the other variables in the model stay the same. For every one unit increase in gpa, ther is a 0.02 increase in the expected value of apply in the log odds scale, given that all of the other variables stay the same.


```{r}
#provide odds-ratio
exp(coef(m11.2))

```


#For pared, every one unit increase in parental education, the odds of “very likely” applying versus “somewhat likely” or “unlikely” applying combined are 2.85 greater, given that all of the other variables in the model are held constant. Also, the odds “very likely” or “somewhat likely” applying versus “unlikely” applying is 2.85 times greater, given that all of the other variables in the model are held constant. For gpa, when a student’s gpa moves 1 unit, the odds of moving from “unlikely” applying to “somewhat likely” or “very likley” applying (or from the lower and middle categories to the high category) are increased by 1.83 (83%).

