---
title: "Individual Analysis 2 Chapter 10"
author: "Denis Kozhokar"
date: "11/11/2018"
output: html_document
---
The dataset that is being used is a very small part of my thesis. An EEG was used to measure engagement and workload (through formulas) to predict errors of omission in a dual task.  The current dataset is from 1 subject, with only the workload index. In addition, each possible trial (Total of 480) was parsed out in 200 sections (These are the predictor variables. I only inported the last 3 predictor variables. The predictor variables are already centered through the forumlas thagt were used. The purpose of this analysis was to use a logistic regression to look at the probability that an error would occur.

```{r}
library(readr)
library(rethinking)
library(rstan)

#import dataset
High_Workload_1016_TaskLoadIndex <- read_excel("Documents/MATLAB/High_Workload_1016_TaskLoadIndex.xlsx")

as.()
as.matrix.data.frame(High_Workload_1016_TaskLoadIndex)
head(High_Workload_1016_TaskLoadIndex)

m10.1 <- map(
    alist(
        admit ~ dbinom( 1 , p ) ,
        logit(p) <- a ,
        a ~ dnorm(0,1)
),
    data = mydata )
precis(m10.1)


m10.3 <- map(
    alist(
        pulled_left ~ dbinom( 1 , p ) ,
        logit(p) <- a + bp*Time1 + bc*Time2 + bd*Time3 ,
        a ~ dnorm(0,10) ,
        bp ~ dnorm(0,1) ,
        bc ~ dnorm(0,1) ,
        bd ~ dnorm(0,1)
), data=High_Workload_1016_TaskLoadIndex)
precis(m10.1)
```


Li ∼ Binomial(1, pi)
logit(pi) = α + (βP + βPCCi)Pi
α ∼ Normal(0, 1) 
β ∼ Normal(0, 1)
σ ∼ Uniform(0, 1)

```{r}

```


