4E1. In the model definition below, which line is the likelihood? 
yi ∼ Normal(μ, σ)

4E2. In the model definition just above, how many parameters are in the posterior distribution?
2

4E3.Using the model definition above, write down the appropriate form of Bayes’ theorem that
includes the proper likelihood and priors.
Pr(μ, σ|h) = ∏ Normal(hi|μ,σ)Normal(μ|0,10)Uniform(σ|0,10)/∏i Normal(hi|μ, σ)Normal(μ|0, 10)Uniform(σ|0, 10)dμdσ

4E4. In the model definition below, which line is the linear model?
μi =α+βxi

4E5. In the model definition just above, how many parameters are in the posterior distribution?
3

4M1. For the model definition below, simulate observed heights from the prior (not the posterior). 
yi ∼ Normal(μ, σ)
μ ∼ Normal(0, 10) 
σ ∼ Uniform(0, 10)

```{r}
library(rethinking)
library(rstan)
library(ggplot2)

mu <- rnorm( 1e4 , 0 , 10 )
sigma <- runif( 1e4 , 0 , 10 )
prior <- rnorm( 1e4 , mu , sigma )
```


4M2. Translate the model just above into a map formula.
```{r}
map_formula <- alist(
  y ~ dnorm( mu , sigma ),
  mu ~ dnorm( 0 , 10 ),
  sigma ~ dunif( 0 , 10 )
)
```


4M3. Translate the map model formula below into a mathematical model definition.
flist <- alist(
    y ~ dnorm( mu , sigma ),
    mu <- a + b*x,
    a ~ dnorm( 0 , 50 ),
    b ~ dunif( 0 , 10 ),
    sigma ~ dunif( 0 , 50 )
)

hi ∼ Normal(μi, σ)
μi = α + βxi
α ∼ Normal(0, 50) 
β ∼ Normal(0, 10)
σ ∼ Uniform(0, 50)

4M4. A sample of students is measured for height each year for 3 years. After the third year,you want to fit a linear regression predicting height using year as a predictor. Write down the mathematical model definition for this regression, using any variable names and priors you choose. Be prepared to defend your choice of priors.
hi ∼ Normal(μi, σ)
μi =α+βxi
α ∼ Normal(165, 20)
β ∼ Normal(0, 10)
σ ∼ Uniform(0, 20)

I used 165 as it is close to the overall mean of height for an individual. In addition, 165 plus or minus 40 encompasses a large portion of human heights. In addition, the standard deviation stays the same, a standard deviation of 20cm would imply that 95% of student heights lie within 0 - 20cm of the average height. the 10 for slope or β  is because we assume that there is a positive slope. Demonstrating that there can be no growth (0) by an individual and adding 10 cm encompasses a large porpotion. The no growth is because many students may have already grown to their maximum height. In addition, most students do not shrink within 3 years.

4M5. Now suppose I tell you that the average height in the first year was 120 cm and that every student got taller each year. Does this information lead you to change your choice of priors? How?
Yes, because at this point, I can  change the confidence metric or α to 120 cm as opposed to 160cm. In addition, we don't need to use a flat prior for the beta coefficient (not use 0)


4M6. Now suppose I tell you that the variance among heights for students of the same age is never more than 64cm. How does this lead you to revise your priors?
I would change the variance or standard deviation to 64. The book mentions that you want leave room for a wide range of variance.






